---
title: "Part A-Exponential Distribution vs Central Limit Theorem"
author: "A Lopez"
date: "20 October 2015"
output: pdf_document
---
Objective:
The aim of this project is to explore the exponential distribution in R and compare it with the Central Limit Theorem. For this project Lambda is set to 0.2 for a thousand simulations. The project considers the distribution of averages of 40 exponentials.

# 1 - Show the sample mean and compare it to the theoretical mean of the distribution

Simulations and Plot:

```{r, echo=TRUE}
# Data: Parameters for the Exponential Distribution
set.seed(900) # Set the seed for random number generation
lambda <- 0.2   # parameter of the distribution, rate parameter
N_sim <- 1:1000 # Number of simulations
N_exp <- 40     # 40 exponentials
#Computations of theoretical and sample mean
means <- data.frame(c = sapply(N_sim, function(c) {mean(rexp(N_exp, lambda))})) #1000 simulations
tMean <- round(1/lambda,4) # mean of the exponential distribution rounded to 4 digits
sample_mean <- round(mean(means$c),4) #sample mean rounded to 4 digits
```
Fig 1 depicts a histogram of the samples and plots both the theoretical and the sample mean. 
```{r fig.width=5, fig.height=4,echo=FALSE}
hist(means$c, col = "grey", main=" Fig.1 Means Distributions, 1,000 Samples", xlab="Mean")
par(srt=90)
abline(v = sample_mean, col = "black", lwd = 2)
text(sample_mean, 100, pos=3, col = "black", "Sample Mean =5.001")
abline(v = 1/lambda, col = "red", lwd = 2)
text(tMean, 100, pos=4, col = "red", "T Mean =5.000" )
```

Analysis: The Sample Mean versus the Theoretical Mean

A) Let $\mu$ be the expected theoretical mean, given by:
$\mu= \frac{1}{\lambda}$ 

replacing values we can calculate $\mu:$ 
```{r, echo=FALSE}
tMean;
```
Let $\bar X$ be the average sample mean of 1000 simulations of 40 randomly sampled exponential distributions.
Computing the mean of the samples we obtain (see code above)
$\bar X:$ 
```{r, echo=FALSE}
sample_mean;
```

Conclusion: Note that the plot and the calculation shown above verify that the values of the expected and the sample mean are very similar with a negligible difference, proving that the simulations are very accurate.

# 2 Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution. 

Expected Values:

Let the std deviation of a exponential distribution of rate $\lambda$ be $\sigma$.
Therefore: 
$\sigma = \frac{1/\lambda}{\sqrt{n}}$

$\sigma:$
```{r, echo=FALSE}
#Computing value
sd <- round((1/lambda)/sqrt(40),4);  sd; # Expected Standard Deviation 
```
And let the variance $Var$ of standard deviation $\sigma$ be
$Var = \sigma^2$

$Var:$
```{r, echo=FALSE}
#Computing value
Var <- sd^2; Var;
```

Sample Calculations:

Let $Var_x$ be the variance of the average sample mean of 1000 simulations of 40 randomly sampled exponential distribution, and $sd_x$ the corresponding standard deviation.
# Variance of the simulations:

```{r, echo=TRUE}
sd_x<-round(sd(means$c),4)# standard deviation
sd_x;
var_x<- round(var(means$c),4)
var_x;
```

Analysis: The standard deviation values are very close, and since the variance is the square of the standard deviations, a negligible difference is expected, demonstrating again the accuracy of the simulations.

# 3- Show that the distributionis approximately normal

Focus on the difference between the distributions of a large collection of 
random exponentials and the distribution of a large collection of averages of 40 exponentials. 


```{r,echo=FALSE}
require(ggplot2)
```

```{r fig.width=3, fig.height=3,echo=FALSE}
ggplot(data = means, aes(x = means[,1])) + 
  geom_histogram(aes(y=..density..), fill = I('grey'), 
                 binwidth = 0.20, color = I('black')) +
  labs(title="40 exponentials-1000 simmulations", x="Means", y="Density")+
  stat_function(fun = dnorm, arg = list(mean = sample_mean, sd = sd_x),color = "blue")+
  stat_function(fun = dnorm, arg = list(mean = tMean, sd = sd), color = "red")+
  geom_vline(xintercept=sample_mean, color="blue")+
  geom_vline(xintercept=tMean, color="blue")

N_sim <- 1:200000 # Increasing Number of simulations
means <- data.frame(c = sapply(N_sim, function(c) {mean(rexp(N_exp, lambda))})) #1000 simulations
sample_mean <- round(mean(means$c),4) #sample mean rounded to 4 digits

ggplot(data = means, aes(x = means[,1])) + 
  geom_histogram(aes(y=..density..), fill = I('grey'), 
                 binwidth = 0.20, color = I('black')) +
  labs(title="40 exponentials-200000 simmulations", x="Means", y="Density")+
  stat_function(fun = dnorm, arg = list(mean = sample_mean, sd = sd_x),color = "blue")+
  stat_function(fun = dnorm, arg = list(mean = tMean, sd = sd), color = "red")+
  geom_vline(xintercept=sample_mean, color="blue")+
  geom_vline(xintercept=tMean, color="blue")
```

Analysis: For lambda=0.2 ,the sample distribution is approximately normal with Mean and Variance values matching the Theoretical mean and variance with a very accurate approximation. 
To further elaborate, Should we increase the sample size from 1000 to x, eg 200000, the distribution of means shape a bell curve of a normal distribution, therefore concluding that it is approximate normal distribution.


#CONCLUSION

Summary Results for Q1, Q2, and Q3
```{r, echo=FALSE}
B = matrix( c(tMean, sd, Var, sample_mean, sd_x, var_x),nrow=3,ncol=2) 
colnames(B) <- c("Expected Value","Simulation")
rownames(B) <- c("Mean", "Std Dev","Variance")
B;
```

1- We see above the distribution is centered, with the theoretical mean vs sample mean presenting a negligible difference. The simulation method is accurate and yields only a difference of .0001 for a 1000 simulations. 
2- The theoretical variance is 0.637 and the sample variance 0.6250, also proving a high degree of accuracy using the sample std deviation and variance.    
3- The graph above shows the distribution approximately normal for 1000 samples and continues the same shape for higher numberof samples,demonstrating again a consistent tendency and approximation to normal distribution.   


 
